### 2021-01-08 paper/paper.tex
adjust algorithm details, finish proposed method section, start results

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.4 (15/5/16)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com) with extensive modifications by
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside,twocolumn]{article}

\usepackage{blindtext} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[english]{babel} % Language hyphenation and typographical rules

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\Roman{subsection}} % roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
\posttitle{\end{center}} % Article title closing formatting
\title{ISODS: An ISODATA-Based Initial Centroid Algorithm} % Article title
\author{%
  \textsc{
    Noah Gardner$^1$\thanks{ngardn10@students.kennesaw.edu}, 
    Sahar Yarmohammadtoosky$^2$\thanks{yarmohamadishr@gmail.com},
    }\\[1ex]
  \textsc{
    Anthony Phan$^1$\thanks{aphan5@students.kennesaw.edu},
    Chih-Cheng Hung$^1$\thanks{chung1@students.kennesaw.edu}} \\[1ex]
  \normalsize $^1$Kennesaw State University \\
  \normalsize College of Computing and Software Engineering \\
  \normalsize Marietta, GA, USA \\[1ex]
  \normalsize $^2$Iran University of Science and Technology \\
  \normalsize Tehran, Iran
}
\date{} % Leave empty to omit a date

\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent 
Iterative Self-Organizing Data Analysis Technique (ISODATA) is a clustering
algorithm based on K-Means that intends to find the correct number of clusters
through merging and splitting. Compared to K-Means, the ISODATA algorithm should
be resilient to initial cluster centers. However, ISODATA is also sensitive to
initial cluster centers which impacts the performance of the algorithm. This
work proposes using the ISODATA approach of splitting clusters to select quality
initial cluster centers. The proposed approach is evaluated on several benchmark
UCI datasets as well as the Salinas-A hyperspectral image dataset. The results
are evaluated using standard clustering evaluation parameters. \\ \\
\textit{\textbf{Keywords:} Clustering, ISODATA, Initial Centroid, Remote Sensing} \\
\end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

% Print the title
\maketitle

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}

\lettrine[nindent=0em,lines=3]{R} emote sensing technology is used when
the data collection takes place at a remote distance from the subject matter,
such as the study of daily weather and long-term climate change, land cover
monitoring, ecosystem modeling of vegetation, and so on. Multispectral imaging
technology is one of the instruments for remote sensing used since the late
1960s. However, multispectral imaging systems use a low number of spectral bands
(three to six spectral bands in different systems). Increasing spectral
resolution in the hyperspectral images (HSI) with several hundred spectral bands
of the observational scene in a single acquisition makes it possible to acquire
detailed examination of observational scenes. The ground-based version of
hyperspectral images has gained immense interest. It has been applied to
numerous electronic imaging applications, including food inspection, forensic
science, medical surgery and diagnosis, and military applications. Because of
the difficulty of obtaining prior knowledge in remote sensing studies,
unsupervised classification and cluster analysis have great importance compared
to supervised classification.

Unsupervised learning is a machine learning technique used when the learning
algorithm has no pre-existing labels and looks for patterns in the data set with
minimum human supervision and without external guidance. In contrast to
supervised learning, where a training set of inputs and observations are given
to the method, unsupervised learning uses only the inputs and should find an
interesting pattern in the data set. Two main approaches are used in
unsupervised learning, named principal component and cluster analysis. In the
clustering method, unlabeled and unsorted data points are grouped and segmented
according to similarities and differences to find structure in the inputs, group
it into meaningful clusters, and understand the data set better. There are
different algorithms for Clustering, such as K-Means, fuzzy C-Means, ISODATA,
and so on. In the K-Means algorithm, the goal is to minimize the sum of the
squared distances points have in the data set from its center. The problem with
the K-Means algorithm is that it requires a priori knowledge about the number of
classes in the data set, and choosing the cluster centers randomly may not
result in optimal partitioning of the data set.

One of the most popular clustering algorithms is called ISODATA clustering,
based on the K-means algorithm with some variations. The ISODATA method's
preference in comparison with the K-means algorithm is that the ISODATA
algorithm uses splitting and merging forms for clusters. The ISODATA method
allows starting from arbitrary cluster centers and obtains better partitioning
results than the K-Means algorithm, but it is not entirely resilient to initial
cluster centers. In this paper, a novel method using the ISODATA approach of
splitting clusters to select high-quality initial cluster centers is proposed.
This new method decreases the number of iterations and improves the overall
performance of the algorithm.

The paper is organized as follows: Section 2 introduces K-Means and ISODATA
clustering algorithms. The new proposed algorithm is explained in Section 3. We
present the experimental results in Section 4. Finally, Section 5 contains
concluding remarks.

%------------------------------------------------

\subsection{K-Means Algorithm}

The K-Means algorithm is an unsupervised clustering algorithm and aims to
partition $N$ observations into $K$ clusters [CITE]. Data points in the same
cluster are very similar to each other and have less variation compared to data
points assigned to other clusters. The user defines $K$, the number of expected
clusters. Following is a description of the K-Means algorithm:

\begin{itemize}
\item \textbf{Step 1:} Choose $K$ cluster centers from the data set.
\item \textbf{Step 2:} Assign each element of the data set to the cluster based
on the minimum distance from the cluster center.
\item \textbf{Step 3:} Re-calculate new cluster centers for each cluster after
the assignment in \textbf{Step 2} by computing the mean of all patterns
assigned to each cluster.
\item \textbf{Step 4:} Repeat \textbf{Step 2 and Step 3} until clustering
convergence occurs, or the set number of iterations is exceeded.
\end{itemize}

For selecting the initial cluster centers in \textbf{Step 1}, there are two
standard methods: Random Centers, and K-Means++ [CITE]. In Random Centers, $K$
samples are randomly chosen from the dataset to serve as the initial cluster
centers. In K-Means++, $1$ sample is randomly chosen from the dataset to serve
as the initial cluster center. Then, $K-1$ samples are iteratively selected,
where the probability of a sample being chosen as a cluster center is the
distance of the sample from previously chosen centers [CITE]. The K-Means
algorithm is susceptible to initial starting cluster centers, where different
initial values can result in a different classification.

%------------------------------------------------

\subsection{ISODATA Algorithm}

ISODATA is another frequently used clustering algorithm and is based on the
K-Means algorithm. Like K-Means, cluster centers are iteratively updated and
reassigned to the samples in the dataset. However, ISODATA determines the number
of clusters dynamically and allows for a different number of cluster centers. In
contrast, the K-Means algorithm requires a fixed user-set parameter of the
number of expected clusters. The advantage found within ISODATA allows for
discovery of a different number of clusters, which is achieved through
discarding clusters with a small number of member samples, and merging and
splitting clusters that meet certain parameters. There are three main ISODATA
parameters:

\begin{itemize}
\item $\theta n$ - minimum number of cluster members, which determines whether
the cluster is discarded,
\item $\theta e$ - maximum standard deviation allowed for a cluster, which
determines whether the cluster is split,
\item $\theta c$ - minimum distance between clusters, which determines whether
two clusters are merged.
\end{itemize}

Additionally, the user must set $K$, the maximum number of clusters to generate.
The ISODATA algorithm is briefly explained below:

\begin{itemize}
\item \textbf{Step 1:} Choose $K$ cluster centers from the data set.
\item \textbf{Step 2:} Assign each element of the dataset to the cluster based
on the minimum distance from the cluster center.
\item \textbf{Step 3:} Re-calculate new cluster centers for each cluster by
computing the mean of all patterns assigned to each cluster.
\item \textbf{Step 4:} Discard clusters that have fewer than ($\theta n$)
members. If any clusters were discarded, move to \textbf{Step 7}.
\item \textbf{Step 5:} If this iteration is an odd numbered iteration or the
number of clusters is less than or equal to  $1/2 * K$, then split clusters
which have a standard deviation in a dimension greater than ($\theta e$). Move
to \textbf{Step 7}.
\item \textbf{Step 6:} If this iteration is an even numbered iteration or the
number of clusters is greater than or equal to $2 * K$, then merge clusters which
have an intercluster distance less than ($\theta c$).
\item \textbf{Step 7:} If this is not the last iteration, start the process
again at step 2. If this is the last iteration or clustering convergence occurs,
then end the process.
\end{itemize}

There are some disadvantages to the ISODATA algorithm. First, ISODATA is
sensitive to the parameters $\theta n$, $\theta e$, and $\theta c$. These
parameters should be fine-tuned for optimal performance based on the attributes
of the dataset being tested. In addition, similar to K-Means, it is sensitive to
initial cluster centers, although less so than K-Means due to the extra
processes of discarding, splitting, and merging. The initial centroids can
affect both the quality of clusters produced and the number of iterations
required to converge [CITE].

%------------------------------------------------

\section{Proposed Method}

The proposed clustering algorithm's objective is to improve the result we obtain
from the ISODATA algorithm by targeting the initial cluster centers. Inspired by
the ISODATA algorithm, Iterative Self-Organizing Data Split (ISODS) uses the
splitting process from ISODATA to create the required initial cluster centers to
use for the ISODATA algorithm. The inputs for the algorithm are the number of
cluster centers $K$ as an integer, and the data to generate centers from. The
algorithm starts by selecting $1$ sample as an initial cluster center. In the
next step, each data is assigned to the cluster and the mean of the cluster is
updated. Then, by the ISODATA splitting process, each cluster is split along the
dimension with the highest standard deviation. The assignments for the new
clusters are calculated based on distance and the centers are updated. This
process is continued until either the number of clusters is greater than or
equal to $K$. If the number of clusters is greater than $K$, the clusters with
the fewest members are discarded one at a time, and the assignments and centers
are updated after each rejection. The algorithm is outlined below:

\begin{itemize}
\item \textbf{Step 1:} Choose $1$ observation as a cluster center from the dataset.
\item \textbf{Step 2:} Assign each element of the dataset to the cluster based
on the minimum distance from the cluster center.
\item \textbf{Step 3:} Re-calculate new cluster centers for each cluster by
computing the mean of all patterns assigned to each cluster.
\item \textbf{Step 4:} If the number of clusters is equal to $K$, then the
process is terminated. If the number of clusters is greater than $K$, then the
clusters with the fewest members is discarded, and the algorithm moves to
\textbf{Step 2}.
\item \textbf{Step 5:} Calculate the standard deviation of each cluster in each
dimension. Then, split each cluster along the dimension with the greatest
standard deviation.
\end{itemize}

This method began as a modified ISODATA process. However, the modification was
unique in that there is a clear distinction from the proposed change and the
beginning of the original process. Although more complex than other methods of
generating initial cluster centers, this approach guarantees variance between
clusters. As a result, ISODS can be used to generate clusters for other
clustering algorithms, such as K-Means. In the next section, ISODS is applied to
both ISODATA and K-Means on some benchmark UCI datasets [CITE] and compared with
Random Centers and K-Means++ Centers.

%------------------------------------------------
\section{Results}

Due to the response of the ISODATA algorithm to the discarding, merging, and
splitting parameters, these parameters should be carefully selected. Since the
goal of this research is to explore the impact of initial cluster centers on
ISODATA, we use a relatively simple approach for generating ISODATA parameters.
Based on the available training data with included labels, each class is
summarized with the mean and standard deviation of that class, as well as the
number of members in each class. To determine $\theta n$, the minimum number of
cluster members used for discarding, we find the class with the fewest members
and use the number of members in that class as $\theta n$. For $\theta e$, the
maximum standard deviation allowed for a cluster, we follow a similar approach
and find the maximum standard deviation within a class. Finally, for $\theta c$,
we calculate the distance between the means of each class, and use the minimum
distance as $\theta c$.

%------------------------------------------------

\section{Conclusion}

\subsection{Subsection One}

A statement requiring citation \cite{Figueredo:2009dg}.
\blindtext % Dummy text

\subsection{Subsection Two}

\blindtext % Dummy text

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
Figueredo, A.~J. and Wolf, P. S.~A. (2009).
\newblock Assortative pairing and life history strategy - a cross-cultural
  study.
\newblock {\em Human Nature}, 20:317--330.

\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{document}