### 2021-01-08 paper/paper.tex
modify style to match conference requirements

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx}
\begin{document}

\title{\bf Patent Duration, Innovative Performance, and Technology Diffusion}
\author{Noah Gardner$^1$,
      Sahar Yarmohammadtoosky$^2$, \\
      Anthony Phan$^3$,
      and Chih-Cheng Hung$^4$\\[1ex]
      \sl \small Kennesaw State University,
      Marietta, GA, USA$^{1, 3, 4}$\\
      \sl \small Iran University of Science and Technology,
      Tehran, Iran$^2$\\[1ex]
      \small $^1$ngardn10@students.kennesaw.edu, 
      \small $^2$yarmohamadishr@gmail.com, \\
      \small $^3$aphan5@students.kennesaw.edu,
      \small $^4$chung1@kennesaw.edu, \\}
\date{}
\maketitle

\begin{abstract}
      \noindent 
      Iterative Self-Organizing Data Analysis Technique (ISODATA) is a clustering
      algorithm based on K-Means that intends to find the correct number of
      clusters through merging and splitting. Compared to K-Means, the ISODATA
      algorithm should be resilient to initial cluster centers. However, ISODATA
      is also sensitive to initial cluster centers which impacts the performance
      of the algorithm. This work proposes using the ISODATA approach of splitting
      clusters to select quality initial cluster centers. The proposed approach is
      evaluated on several benchmark UCI datasets as well as the Salinas-A
      hyperspectral image dataset. The results are evaluated using standard
      clustering evaluation parameters. \\
      {\bf Keywords :} Clustering, ISODATA, Initial Centroid, Remote Sensing
\end{abstract}

% =========================================================================== %

\section{Introduction}

Remote sensing technology is used when the data collection takes place at a
remote distance from the subject matter, such as the study of daily weather and
long-term climate change, land cover monitoring, ecosystem modeling of
vegetation, and so on \cite{Jensen1996,Mallapragada2018}. Multispectral imaging
technology is one of the instruments for remote sensing used since the late
1960s \cite{Khan2018,Landgrebe1999}. However, multispectral imaging systems use
a low number of spectral bands (three to six spectral bands in different
systems). Increasing spectral resolution in the hyperspectral images (HSI) with
several hundred spectral bands of the observational scene in a single
acquisition makes it possible to acquire detailed examination of observational
scenes \cite{Khan2018}. The ground-based version of hyperspectral images has
gained immense interest. It has been applied to numerous electronic imaging
applications, including food inspection, forensic science, medical surgery and
diagnosis, and military applications
\cite{Govender2009,Adam2009,Fischer2006,Carrasco2003}. Because of the difficulty
of obtaining prior knowledge in remote sensing studies, unsupervised
classification and cluster analysis have great importance compared to supervised
classification \cite{Abbas2016}.

Unsupervised learning is a machine learning technique used when the learning
algorithm has no pre-existing labels and looks for patterns in the data set with
minimum human supervision and without external guidance. In contrast to
supervised learning, where a training set of inputs and observations are given
to the method, unsupervised learning uses only the inputs and should find an
interesting pattern in the data set. Two main approaches are used in
unsupervised learning, named principal component \cite{Mallapragada2018} and
cluster analysis \cite{Xu2015}. In the clustering method, unlabeled and unsorted
data points are grouped and segmented according to similarities and differences
to find structure in the inputs, group it into meaningful clusters, and
understand the data set better. There are different algorithms for Clustering,
such as K-Means, fuzzy C-Means, ISODATA \cite{Ball1965}, and so on. In the
K-Means algorithm, the goal is to minimize the sum of the squared distances
points have in the data set from its center. The problem with the K-Means
algorithm is that it requires a priori knowledge about the number of classes in
the data set, and choosing the cluster centers randomly may not result in
optimal partitioning of the data set.

One of the most popular clustering algorithms is called ISODATA clustering,
based on the K-means algorithm with some variations. The ISODATA method's
preference in comparison with the K-means algorithm is that the ISODATA
algorithm uses splitting and merging forms for clusters. The ISODATA method
allows starting from arbitrary cluster centers and obtains better partitioning
results than the K-Means algorithm, but it is not entirely resilient to initial
cluster centers. In this paper, a novel method using the ISODATA approach of
splitting clusters to select quality initial cluster centers is proposed.

The paper is organized as follows: Section 2 introduces K-Means and ISODATA
clustering algorithms. The new proposed algorithm is explained in Section 3. We
present the experimental results in Section 4. Finally, Section 5 contains
concluding remarks.

% =========================================================================== %

\section{Background}

\subsection{K-Means Algorithm}

The K-Means algorithm is an unsupervised clustering algorithm and aims to
partition $N$ observations into $K$ clusters [CITE]. Data points in the same
cluster are very similar to each other and have less variation compared to data
points assigned to other clusters. The user defines $K$, the number of expected
clusters. Following is a description of the K-Means algorithm:

\begin{itemize}
      \item \textbf{Step 1:} Choose $K$ cluster centers from the data set.
      \item \textbf{Step 2:} Assign each element of the data set to the cluster
            based on the minimum distance from the cluster center.
      \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster
            after the assignment in \textbf{Step 2} by computing the mean of all
            patterns assigned to each cluster.
      \item \textbf{Step 4:} Repeat \textbf{Step 2 and Step 3} until clustering
            convergence occurs, or the set number of iterations is exceeded.
\end{itemize}

For selecting the initial cluster centers in \textbf{Step 1}, there are two
standard methods: Random Centers, and K-Means++ [CITE]. In Random Centers, $K$
samples are randomly chosen from the dataset to serve as the initial cluster
centers. In K-Means++, $1$ sample is randomly chosen from the dataset to serve
as the initial cluster center. Then, $K-1$ samples are iteratively selected,
where the probability of a sample being chosen as a cluster center is the
distance of the sample from previously chosen centers [CITE]. The K-Means
algorithm is susceptible to initial starting cluster centers, where different
initial values can result in a different classification.

\subsection{ISODATA Algorithm}

ISODATA is another frequently used clustering algorithm and is based on the
K-Means algorithm. Like K-Means, cluster centers are iteratively updated and
reassigned to the samples in the dataset. However, ISODATA determines the number
of clusters dynamically and allows for a different number of cluster centers. In
contrast, the K-Means algorithm requires a fixed user-set parameter of the
number of expected clusters. The advantage found within ISODATA allows for
discovery of a different number of clusters, which is achieved through
discarding clusters with a small number of member samples, and merging and
splitting clusters that meet certain parameters. There are three main ISODATA
parameters:

\begin{itemize}
      \item $\theta n$ - minimum number of cluster members, which determines
            whether the cluster is discarded,
      \item $\theta e$ - maximum standard deviation allowed for a cluster, which
            determines whether the cluster is split,
      \item $\theta c$ - minimum distance between clusters, which determines
            whether two clusters are merged.
\end{itemize}

Additionally, the user must set $K$, the maximum number of clusters to generate.
The ISODATA algorithm is briefly explained below:

\begin{itemize}
      \item \textbf{Step 1:} Choose $K$ cluster centers from the data set.
      \item \textbf{Step 2:} Assign each element of the dataset to the cluster
            based on the minimum distance from the cluster center.
      \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster by
            computing the mean of all patterns assigned to each cluster.
      \item \textbf{Step 4:} Discard clusters that have fewer than ($\theta n$)
            members. If any clusters were discarded, move to \textbf{Step 7}.
      \item \textbf{Step 5:} If this iteration is an odd numbered iteration or the
            number of clusters is less than or equal to  $1/2 * K$, then split
            clusters which have a standard deviation in a dimension greater than
            ($\theta e$). Move to \textbf{Step 7}.
      \item \textbf{Step 6:} If this iteration is an even numbered iteration or
            the number of clusters is greater than or equal to $2 * K$, then merge
            clusters which have an intercluster distance less than ($\theta c$).
      \item \textbf{Step 7:} If this is not the last iteration, start the process
            again at step 2. If this is the last iteration or clustering
            convergence occurs, then end the process.
\end{itemize}

There are some disadvantages to the ISODATA algorithm. First, ISODATA is
sensitive to the parameters $\theta n$, $\theta e$, and $\theta c$. These
parameters should be fine-tuned for optimal performance based on the attributes
of the dataset being tested. In addition, similar to K-Means, it is sensitive to
initial cluster centers, although less so than K-Means due to the extra
processes of discarding, splitting, and merging. The initial centroids can
affect both the quality of clusters produced and the number of iterations
required to converge [CITE].

% =========================================================================== %

\section{Proposed Method}

The proposed clustering algorithm's objective is to improve the result we obtain
from the ISODATA algorithm by targeting the initial cluster centers. Inspired by
the ISODATA algorithm, Iterative Self-Organizing Data Split (ISODS) uses the
splitting process from ISODATA to create the required initial cluster centers to
use for the ISODATA algorithm. The inputs for the algorithm are the number of
cluster centers $K$ as an integer, and the data to generate centers from. The
algorithm starts by selecting $1$ sample as an initial cluster center. In the
next step, each data is assigned to the cluster and the mean of the cluster is
updated. Then, by the ISODATA splitting process, each cluster is split along the
dimension with the highest standard deviation. The assignments for the new
clusters are calculated based on distance and the centers are updated. This
process is continued until either the number of clusters is greater than or
equal to $K$. If the number of clusters is greater than $K$, the clusters with
the fewest members are discarded one at a time, and the assignments and centers
are updated after each rejection. The algorithm is outlined below:

\begin{itemize}
      \item \textbf{Step 1:} Choose $1$ observation as a cluster center from the
            dataset.
      \item \textbf{Step 2:} Assign each element of the dataset to the cluster
            based on the minimum distance from the cluster center.
      \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster by
            computing the mean of all patterns assigned to each cluster.
      \item \textbf{Step 4:} If the number of clusters is equal to $K$, then the
            process is terminated. If the number of clusters is greater than $K$,
            then the clusters with the fewest members is discarded, and the
            algorithm moves to \textbf{Step 2}.
      \item \textbf{Step 5:} Calculate the standard deviation of each cluster in
            each dimension. Then, split each cluster along the dimension with the
            greatest standard deviation.
\end{itemize}

This method began as a modified ISODATA process. However, the modification was
unique in that there is a clear distinction from the proposed change and the
beginning of the original process. Although more complex than other methods of
generating initial cluster centers, this approach guarantees variance between
clusters. As a result, ISODS can be used to generate clusters for other
clustering algorithms, such as K-Means.

% =========================================================================== %

\section{Results}

Due to the response of the ISODATA algorithm to the discarding, merging, and
splitting parameters, these parameters should be carefully selected. Since the
goal of this research is to explore the impact of initial cluster centers on
ISODATA, we use a relatively simple approach for generating ISODATA parameters.
Based on the available training data with included labels, each class is
summarized with the mean and standard deviation of that class, as well as the
number of members in each class. Typically, since we are working with
unsupervised learning algorithms, class information may not be available.

To determine $\theta n$, the minimum number of cluster members used for
discarding, we find the class with the fewest members and use the number of
members in that class as $\theta n$. For $\theta e$, the maximum standard
deviation allowed for a cluster, we follow a similar approach and find the
maximum standard deviation within a class. Finally, for $\theta c$, we calculate
the distance between the means of each class, and use the minimum distance as
$\theta c$.

For experiments 1 and 2, we explore applying ISODS to ISODATA and K-Means by
using benchmark UCI datasets Iris, Wine, and Glass. We also show results from
applying Random Centers and K-Means++ Centers. To evaluate the cluster results,
external cluster validation measures are used. We utilize pair-counting measures
which rely on counting matching pairs between two sets of data to determine
agreement and disagreement \cite{Rezaei2016}. The pair-counting measures applied
are \textit{rand index} (RI - agreement probability), \textit{adjusted rand
      index} (ARI), \textit{mirkin's index} (MI - disagreement probability), and
\textit{hubert's index} (HI - $P(agree) - P(disagree)$)\cite{Hubert1985,
      Meila2003}.

% =========================================================================== %

\section{Conclusion}
Put Conclusion here.

% =========================================================================== %

\subsection*{Acknowledgements}
Put Acknowledgements here.

\vspace{5mm}

\begin{center}
      {\large \bf References}
\end{center}

\bibliographystyle{unsrt}
\bibliography{paper}

\end{document}